---
title: "p8105_hw3_wz2630"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

# library with datasets used in this homework
library(p8105.datasets)

# figure settings
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
```{r}
data("instacart")
```

There are 1,384,617 observations and 15 variables in the 'instacart' dataset. 4 of 15 are character data type and the remaining is integer data type. Key variables include order_id, product_id, user_id, product_name, aisle_id, department_id, not including all.
# what are illustrative examples?


```{r}
# How many aisles are there, and which aisles are the most items ordered from?
aisle_count = 
  instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```
There are 134 aisles. Fresh vegetables (aisle_id = 83) and fresh fruits (aisle_id = 24) have the most items ordered. Fresh vegetables aisle have 150,609 items ordered and fresh fruits aisle have 150,473 items ordered.


```{r}
# Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> #fct_reorder turns aisle into factor variables based on count, n
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, color = "navy"))
  
```

```{r}
# Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table

instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  # min_rank gives the same rank to ties based on product_names we counted; ranking function
  filter(rank < 4) |> 
  # we want top 3 products
  arrange(desc(n)) |>
  knitr::kable() # returns a table
```

```{r}
# Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |> 
  summarise(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |> 
  knitr::kable(digits = 2)

```


## Problem 2
```{r, message=FALSE}
data("brfss_smart2010")
```

Data cleaning
```{r}
#brfss_df = 
  #brfss_smart2010 |> 
  #janitor::clean_names() |> 
  #filter(response = c("Excellent", "Very good", "Good", "Fair", "Poor"))
```



## Problem 3
