---
title: "p8105_hw3_wz2630"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

#extra color palette:
library(wesanderson)

# library with datasets used in this homework
library(p8105.datasets)

# figure settings
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
```{r}
data("instacart")
```

There are 1,384,617 observations and 15 variables in the 'instacart' dataset. 4 of 15 are character data type and the remaining is integer data type. Key variables include order_id, product_id, user_id, product_name, aisle_id, department_id, not including all.
# what are illustrative examples?


```{r}
# How many aisles are there, and which aisles are the most items ordered from?
aisle_count = 
  instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```
There are 134 aisles. Fresh vegetables (aisle_id = 83) and fresh fruits (aisle_id = 24) have the most items ordered. Fresh vegetables aisle have 150,609 items ordered and fresh fruits aisle have 150,473 items ordered.


```{r}
# Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> #fct_reorder turns aisle into factor variables based on count, n
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, color = "navy"))
  
```

```{r}
# Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table

instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  # min_rank gives the same rank to ties based on product_names we counted; ranking function
  filter(rank < 4) |> 
  # we want top 3 products
  arrange(desc(n)) |>
  knitr::kable() # returns a table
```

```{r}
# Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |> 
  summarise(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |> 
  knitr::kable(digits = 2)

```


## Problem 2
```{r}
data("brfss_smart2010")
```

Data cleaning
```{r}
brfss = 
  brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"), topic == "Overall Health") |> 
  mutate(response = factor(response, 
                           levels = c("Excellent", "Very good", "Good", "Fair", "Poor", ordered = TRUE)))
```

In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss |> 
  filter(year == 2002) |> 
  group_by(locationabbr) |> 
  summarise(locations_count = n_distinct(locationdesc)) |> 
  filter(locations_count >= 7)

brfss |> 
  filter(year == 2010) |> 
  group_by(locationabbr) |> 
  summarise(locations_count = n_distinct(locationdesc)) |> 
  filter(locations_count >= 7)
```
In 2002, **6** states were observed at 7 or more locations. Those states are **CT, FL, MA, NC, NJ, PA**.

In 2010, **14** states were observed at 7 or more locations. Those states are **CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA**.


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
excellent_responses = 
  brfss |> 
  filter(response == "Excellent") |>
  group_by(year, locationabbr) |> 
  summarise(data_value_avg = mean(data_value))

# spaghetti plot:
excellent_responses |> 
  rename("State" = locationabbr) |> 
  ggplot(aes(x = year, y = data_value_avg, color = State)) + 
  geom_line(alpha = .5) +
  geom_point() + 
  labs(x = "Year", 
       y = "Average Data Values", 
       title = "Average data values across years for each state") + 
  theme(legend.position = "right",
        legend.box.background = element_rect())
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State
```{r}
# two_panel_plot
brfss |> 
  filter(year %in% c(2006, 2010), locationabbr == "NY") |> 
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_boxplot() + 
  facet_wrap(~year) + 
  labs(x = "Response", 
       y = "Data Value",
       title = "Distribution of responses among locations in NY State (2006 and 2010)") + 
  theme(legend.position = "none",
        panel.spacing = unit(1, "lines"))
```


## Problem 3

load data and clean
```{r}
covar = 
  read_csv(file = "./data/nhanes_covar.csv",
           skip = 4) |> 
  janitor::clean_names() |> 
  filter(age >= 21) |> 
  drop_na()

accel = 
  read_csv(file = "./data/nhanes_accel.csv") |> 
  janitor::clean_names()
  
```

Merger demographic and accelerometer datasets:
```{r}
nhanes = 
  inner_join(covar, accel, by = "seqn")

```

Factor variables and rename
```{r}
nhanes$sex = factor(nhanes$sex, 
                                  levels = c("1", "2"),
                                  labels = c("male", "female"))

nhanes$education = factor(nhanes$education, 
                                  levels = c("1", "2", "3"),
                                  labels = c("less than high school", "high school equivalent", "more than high school"))
```

Encode data with reasonable variable classes
```{r}
nhanes$seqn = as.integer(nhanes$seqn)
nhanes$age = as.integer(nhanes$age)
nhanes$bmi = as.integer(nhanes$bmi)
```


Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

Table of number of men and women in each education category
```{r}
nhanes |> 
  group_by(sex, education) |> 
  summarize(count = n()) |> 
  pivot_wider(names_from = education, values_from = count) |> 
  knitr::kable(digits = 1)
```
__Comments:__
The table shows that there are about the same males(27) and females(28) in the "less than high school" category. There are more male(35) than female(23) in the "high school equivalent" category. There are also about the same male(56) and female(59) in the "more than high school" education category.


A visualization of the age distributions for men and women in each education category
```{r}
nhanes |> 
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot(alpha = 0.50) +
  labs(x = "Education level",
       y = "Age",
       title = "Age distributions for men and women in each education category") +
  theme(legend.position = "right",
        legend.box.background = element_rect()) + #gives legend a box
  scale_fill_manual(values = wes_palette("Moonrise3", n = 2)) #extra: changes color of bp
```
__Comments:__
Based on the boxplots, we can see that the age distributions for "less than high school" and "more than high school" categories are similar when compared between male and females in their respective groups. However, the age distributions for "high school equivalent" category are quite different between male and females; males with "high school equivalent" are younger than females overall from the same category. This category also has the largest age range.


total activity over the day. 
Using your tidied dataset, 
aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.
```{r}
# create new variable
nhanes_total <-
  nhanes |> 
    mutate(total_activity = rowSums(across(min1:min1440), na.rm = TRUE))

# make plots
nhanes_total |> 
ggplot(aes(x = age, y = total_activity, color = sex)) + 
  geom_point(alpha = 0.5, size = 1.5) + 
  geom_smooth(method = lm, se = FALSE) + 
  facet_wrap(~education) + 
  labs(x = "Age",
       y = "Total Activity in a Day",
       title = "Total Minutes of Activity in a Day Grouped by Education Level (Male vs Female)") +
  theme(legend.position = "right",
        legend.box.background = element_rect(),
        panel.spacing = unit(1, "lines")) +
  scale_fill_manual(values = wes_palette("Moonrise3", n = 2))

```
**Comment:**
Overall, there is a negative association between age and total physical activity across all 3 education levels. There is a decreased amount of physical activity in a day as age increases. 
The "less than high school" education level has overall more physical activity. Females tend to have more physical activity for both "high school equivalent" and "more than high school" education levels.


Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. 
```{r}
# first we want to pivot the observation so each minute is its own observation and we can plot these xalues as x
nhanes_pivotlonger <-
  nhanes |> 
  pivot_longer(cols = starts_with("min"), 
               names_to = "minute", #turned "min_" into a minute column
               values_to = "activity")  #turned all the time values to activity column
  
nhanes_pivotlonger |> 
  ggplot(aes(x = minute, y = activity, color = sex)) + 
  geom_point(alpha = 0.1, size = 0.5) + 
  geom_smooth(se = FALSE) + 
  facet_wrap(~education) + 
  labs(x = "Minutes (in 24 hours)",
       y = "Activity time (minutes)",
       title = "24-hour activity time courses for each education level") + 
  theme(legend.position = "right",
        legend.box.background = element_rect(),
        panel.spacing = unit(1, "lines"))
```
**Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences:**

- Based on the graphs, it is difficult to differentiate whether there is difference in amount of physical activty in a 24hr period between male and female.
- 











